---
title: "HighFreq_Trained_Models"
output: html_document
date: "2022-10-13"
---
This file sets up the tuning/predictor set comparisons for linear regression, lasso regression, and random forests for Richland-Chambers only.

```{r setup}
library(tidymodels)
library(here)
library(lubridate)
library(tidyverse)
library(doParallel)



#Data were created in Collate_ShortTermANDValidation_Data.R - read XXX.csv
#RC_train <- read_csv("RC_pred_highfreq_training.csv")|>
 # drop_na(Depth_m,Cloud_cov_fraction,DewPoint_C,AirTemp_C,Pressure_hPa, WWind_m.s,SWind_m.s)
  #mutate(hour_of_day = as.numeric(hour(time_of_day)), .keep = "unused")|>
 # select(-profileId)


load(here("Predictive_Models/Create_Recipes_Environment.Rdata"))

```



# Simple linear model - high-freq 
```{r select best lm}
#set engine - very easy in this case
lm_spec <- linear_reg() |> 
  set_engine("lm")

lm_models <- workflow_set(
  preproc = recipes_list,
  models = list(glm = lm_spec))


cl <- makePSOCKcluster(14)
registerDoParallel(cl)
lm_rsmp <- lm_models|>
  workflow_map("fit_resamples", verbose = TRUE, resamples = vfold_cv(rec_training, v = 10, repeats = 5, strata = Depth_m))
  
lm_metrics<-collect_metrics(lm_rsmp)

lm_plot <- lm_metrics|>
  mutate(.metric = fct_recode(.metric, `R-squared` = "rsq", RMSE = "rmse"))|>
  ggplot(aes(x = wflow_id, y = mean))+
  geom_point()+
  geom_errorbar(aes(ymin = mean-(1.96*std_err), ymax = mean+(1.96*std_err)))+
  facet_wrap(.~.metric, scales = "free")+
  xlab("Model ID")+
  ylab("Metric Score")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
lm_plot


```


Want to save the best lm from rmse for later use on validation data
```{r save best lm}
best_lm_id<-lm_rsmp |> 
  rank_results(rank_metric = "rmse")|>
  slice_head()|>
  pull(wflow_id)
best_lm_id
  

best_lm_fit <- lm_rsmp|>
  extract_workflow(best_lm_id)|>
  fit(data = rec_training)

saveRDS(best_lm_fit, "trained_workflows/lm_highfreq_fit.rds")

```

recipe_list <- list("lag0" = rec_lag0,"lag1" = rec_lag1,"lag6" = rec_lag6,"lag12" = rec_lag12,"lag24" = rec_lag24,"lag36" = rec_lag36,"lag48" = rec_lag48,"lag72" = rec_lag72,"lag144" = rec_lag144, 
                    "lag0_int" = rec_lag0_int,"lag1_int" = rec_lag1_int,"lag6_int" = rec_lag6_int,"lag12_int" = rec_lag12_int,"lag24_int" = rec_lag24_int,"lag36_int" = rec_lag36_int,"lag48_int" = rec_lag48_int,"lag72_int" = rec_lag72_int,"lag144_int" = rec_lag144_int)

# Lasso
```{r}
lambda_grid <- grid_regular(penalty(), levels = 50) #column of penalties evenly spread from 1e-10 to 1

tune_spec <- linear_reg(penalty = tune(), mixture = 1)|>
  set_engine("glmnet")


lasso_models <- workflow_set(
  preproc = recipes_list,
  models = list(lasso = tune_spec))

cl <- makePSOCKcluster(14)
registerDoParallel(cl)
lasso_tune <- lasso_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(rec_training, v = 10, repeats = 5, strata = Depth_m), grid = lambda_grid)

#saveRDS(lasso_tune, "trained_workflows/lasso_tune_object.rds")
#lasso_tune <- readRDS("trained_workflows/lasso_tune_object.rds")

  autoplot(lasso_tune, select_best=TRUE, rank_metric = "rmse", metric = "rmse")+
   geom_text(aes(y = mean - 0.1, label = wflow_id), angle = 45) +
   lims(y = c(1, 3)) +
   theme(legend.position = "none")
  

#this will get the best resulting lasso model from whichever recipe appeared to perform best - autoplot looks like lag0
lasso_best_results <- lasso_tune|>collect_metrics()|>
  group_by(.metric,.config)|>arrange(mean, .by_group = TRUE)|>group_by(wflow_id)|>filter(.config == pluck(.config[1]))

lasso_predselect_plot <- lasso_best_results|>
  mutate(.metric = fct_recode(.metric, `R-squared` = "rsq", RMSE = "rmse"),
         wflow_id = as_factor(wflow_id),
         wflow_id = fct_reorder2(wflow_id, desc(.metric), desc(mean)))|>
  ggplot(aes(x = wflow_id, y = mean))+
  geom_point()+
  geom_errorbar(aes(ymin = mean-(1.96*std_err), ymax = mean+(1.96*std_err)))+
  facet_wrap(.~.metric, scales = "free")+
  xlab("Model ID")+
  ylab("Metric Score")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

lasso_predselect_plot

lasso_best <-lasso_tune |> 
  extract_workflow_set_result("intlag0_lasso")|>
  select_best(metric = "rmse")

lasso_final <- lasso_tune|>
  extract_workflow("intlag0_lasso")|>
  finalize_workflow(lasso_best)

#tuning parameter plot
lasso_tune |> 
  extract_workflow_set_result("intlag0_lasso")|>
  collect_metrics()|>
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_errorbar(aes(
    ymin = mean - std_err,
    ymax = mean + std_err
  ),
  alpha = 0.5
  ) +
  geom_line(size = 1.5) +
  facet_wrap(~.metric, scales = "free", nrow = 2) +
  scale_x_log10() +
  theme(legend.position = "none")

saveRDS(lasso_final, "trained_workflows/lasso_final.rds")

```





# Random Forest 
```{r randfor mods}
tune_randfor <- rand_forest(
  mtry = tune(),
  trees = 500, 
  min_n = tune()) |>
  set_mode("regression") %>%
  set_engine("ranger", importance = "impurity")


########################### Workflow Sets
randfor_lag0_models <- workflow_set(
  preproc = list(lag0 = rec_lag0),
  models = list(randfor = tune_randfor))

randfor_lag2_models <- workflow_set(
  preproc = list(lag2 = rec_lag2),
  models = list(randfor = tune_randfor))

randfor_lag6_models <- workflow_set(
  preproc = list(lag6 = rec_lag6),
  models = list(randfor = tune_randfor))

randfor_lag12_models <- workflow_set(
  preproc = list(lag12 = rec_lag12),
  models = list(randfor = tune_randfor))

randfor_lag24_models <- workflow_set(
  preproc = list(lag24 = rec_lag24),
  models = list(randfor = tune_randfor))



randfor_lag48_models <- workflow_set(
  preproc = list(lag48 = rec_lag48),
  models = list(randfor = tune_randfor))

randfor_lag72_models <- workflow_set(
  preproc = list(lag72 = rec_lag72),
  models = list(randfor = tune_randfor))


randfor_intlag0_models <- workflow_set(
  preproc = list(intlag0 = rec_int_lag0),
  models = list(randfor = tune_randfor))

randfor_intlag2_models <- workflow_set(
  preproc = list(intlag2 = rec_int_lag2),
  models = list(randfor = tune_randfor))

randfor_intlag6_models <- workflow_set(
  preproc = list(intlag6 = rec_int_lag6),
  models = list(randfor = tune_randfor))

randfor_intlag12_models <- workflow_set(
  preproc = list(intlag12 = rec_int_lag12),
  models = list(randfor = tune_randfor))

randfor_intlag24_models <- workflow_set(
  preproc = list(intlag24 = rec_int_lag24),
  models = list(randfor = tune_randfor))


randfor_intlag48_models <- workflow_set(
  preproc = list(intlag48 = rec_int_lag48),
  models = list(randfor = tune_randfor))

randfor_intlag72_models <- workflow_set(
  preproc = list(intlag72 = rec_int_lag72),
  models = list(randfor = tune_randfor))



### tune models
cl <- makePSOCKcluster(14)
doParallel::registerDoParallel(cl)
randfor_lag0_tuned <- randfor_lag0_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag0_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_lag2_tuned <- randfor_lag2_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag2_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_lag6_tuned <- randfor_lag6_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag6_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_lag12_tuned <- randfor_lag12_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag12_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_lag24_tuned <- randfor_lag24_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag24_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_lag48_tuned <- randfor_lag48_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag48_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_lag72_tuned <- randfor_lag72_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag72_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)


doParallel::registerDoParallel(cl)
randfor_intlag0_tuned <- randfor_intlag0_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag0_int_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_intlag2_tuned <- randfor_intlag2_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag2_int_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_intlag6_tuned <- randfor_intlag6_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag6_int_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_intlag12_tuned <- randfor_intlag12_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag12_int_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_intlag24_tuned <- randfor_intlag24_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag24_int_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_intlag48_tuned <- randfor_intlag48_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag48_int_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)
doParallel::registerDoParallel(cl)
randfor_intlag72_tuned <- randfor_intlag72_models|>
  workflow_map(verbose = TRUE, "tune_grid", resamples = vfold_cv(lag72_int_training, v = 10, repeats = 5, strata = Depth_m), grid = 20)


###

randfor_lag0_wf_best <- randfor_lag0_tuned|>extract_workflow_set_result("lag0_randfor")|>collect_metrics()|>
    group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "lag0_randfor")
randfor_lag2_wf_best <- randfor_lag2_tuned|>extract_workflow_set_result("lag2_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "lag2_randfor")
randfor_lag6_wf_best <- randfor_lag6_tuned|>extract_workflow_set_result("lag6_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "lag6_randfor")
randfor_lag12_wf_best <- randfor_lag12_tuned|>extract_workflow_set_result("lag12_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "lag12_randfor")
randfor_lag24_wf_best <- randfor_lag24_tuned|>extract_workflow_set_result("lag24_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "lag24_randfor")
randfor_lag48_wf_best <- randfor_lag48_tuned|>extract_workflow_set_result("lag48_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "lag48_randfor")
randfor_lag72_wf_best <- randfor_lag72_tuned|>extract_workflow_set_result("lag72_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "lag72_randfor")


randfor_intlag0_wf_best <- randfor_intlag0_tuned|>extract_workflow_set_result("intlag0_randfor")|>collect_metrics()|>
    group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "intlag0_randfor")
randfor_intlag2_wf_best <- randfor_intlag2_tuned|>extract_workflow_set_result("intlag2_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "intlag2_randfor")
randfor_intlag6_wf_best <- randfor_intlag6_tuned|>extract_workflow_set_result("intlag6_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "intlag6_randfor")
randfor_intlag12_wf_best <- randfor_intlag12_tuned|>extract_workflow_set_result("intlag12_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "intlag12_randfor")
randfor_intlag24_wf_best <- randfor_intlag24_tuned|>extract_workflow_set_result("intlag24_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "intlag24_randfor")
randfor_intlag48_wf_best <- randfor_intlag48_tuned|>extract_workflow_set_result("intlag48_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "intlag48_randfor")
randfor_intlag72_wf_best <- randfor_intlag72_tuned|>extract_workflow_set_result("intlag72_randfor")|>collect_metrics()|>
  group_by(.metric)|>arrange(mean, .by_group = TRUE)|>ungroup()|>filter(.config == pluck(.config[1]))|>mutate(wflow_id = "intlag72_randfor")

#combine all the *_tuned workflow_map objects and select best
randfor_mods <- list(randfor_lag0_tuned,randfor_lag2_tuned,randfor_lag6_tuned,randfor_lag12_tuned,
     randfor_lag24_tuned,randfor_lag48_tuned,randfor_lag72_tuned,randfor_intlag0_tuned,
     randfor_intlag2_tuned,randfor_intlag6_tuned,randfor_intlag12_tuned,
     randfor_intlag24_tuned,randfor_intlag48_tuned,randfor_intlag72_tuned)

saveRDS(randfor_mods, file = "trained_workflows/randfor_mods.Rds")
saveRDS(randfor_lag0_tuned, file = "trained_workflows/randfor_tuneresults_lag0.Rds")
randfor_lag0_tuned<-readRDS(here("Predictive_Models/trained_workflows/randfor_mod_lag0.Rds"))

randfor_best_results <- bind_rows(randfor_lag0_wf_best,randfor_lag2_wf_best,randfor_lag6_wf_best,randfor_lag12_wf_best,
                                  randfor_lag24_wf_best,randfor_lag48_wf_best,randfor_lag72_wf_best,
                                  randfor_intlag0_wf_best, randfor_intlag2_wf_best,randfor_intlag6_wf_best,
                                  randfor_intlag12_wf_best, randfor_intlag24_wf_best, randfor_intlag48_wf_best)


randfor_predselect_plot <- randfor_best_results|>
  mutate(.metric = fct_recode(.metric, `R-squared` = "rsq", RMSE = "rmse"),
         wflow_id = as_factor(wflow_id),
         wflow_id = fct_reorder2(wflow_id, desc(.metric), desc(mean)))|>
  ggplot(aes(x = wflow_id, y = mean))+
  geom_point()+
  geom_errorbar(aes(ymin = mean-(1.96*std_err), ymax = mean+(1.96*std_err)))+
  facet_wrap(.~.metric, scales = "free")+
  xlab("Model ID")+
  ylab("Metric Score")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
randfor_predselect_plot

randfor_best <- randfor_lag0_tuned |>
  extract_workflow_set_result("lag0_randfor")|>
  select_best(metric = "rmse")
#min_n = 4, mtry = 6
randfor_final <- randfor_lag0_tuned|>
  extract_workflow("lag0_randfor")|>
  finalize_workflow(randfor_best)



final_randfor <- last_fit(randfor_final, lag0_split)
final_vi<-final_randfor|>extract_fit_parsnip()|>vip(num_features = 80)|>pluck("data")
write_csv(final_vi, "trained_workflows/randfor_lag0_vip_scores.csv")
saveRDS(final_randfor, "trained_workflows/randfor_fit.Rds")
#final_randfor <- readRDS("trained_workflows/randfor_fit.Rds")
```

```{r randfor preds}

blah <- final_randfor|>extract_workflow() #can be used to look at min_n and mtry
blah2<-blah |> extract_recipe() #checked that #number features in selected recipe >= mtry...

final_randfor_metrics <- final_randfor |> 
  collect_metrics()
final_randfor_preds <- final_randfor |>
  collect_predictions()|>
  bind_cols(lag0_int_testing|>select(-DO_mg.L))|>
  filter(!is.na(val_period))

final_randfor_preds|>
  filter(Depth_m == 1)|>
ggplot()+
  geom_point(aes(x = Datetime, y = .pred))+
  geom_line(aes(x = Datetime, y = DO_mg.L))+
  ylab("DO (mg/L)")+
  xlab("Datetime")+
  facet_wrap(.~val_period, scales = "free")

```

